---
title: "Reproducible Research: Peer Assessment 1"
author: "LiYen"
date: "September 11, 2015"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r, echo=TRUE}
data <- read.csv("activity.csv", colClasses = c("integer", "Date", "factor"))
data$month <- as.numeric(format(data$date, "%m"))
dataNoNA <- na.omit(data)
rownames(dataNoNA) <- 1:nrow(dataNoNA)
head(dataNoNA)
dim(dataNoNA)
library(ggplot2)
```


## What is mean total number of steps taken per day?
Calculate and report the mean and median of the total number of steps taken per day
```{r, echo=TRUE}
totalSteps <- aggregate(dataNoNA$steps, list(Date = dataNoNA$date), sum)$x
mean(totalSteps)
median(totalSteps)
```

  Make a Histrogram
  
```{r ggplot, echo=TRUE}
ggplot(dataNoNA, aes(date, steps)) + geom_bar(stat = "identity", colour = "blue", fill = "blue", width = 0.7) + facet_grid(. ~ month, scales = "free") + labs(title = "Histogram of Total Number of Steps Taken Each Day", x = "Date", y = "Total number of steps")
```


## What is the average daily activity pattern?
  Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
  
```{r, echo=TRUE}
avgSteps <- aggregate(dataNoNA$steps, list(interval = as.numeric(as.character(dataNoNA$interval))), mean)
names(avgSteps)[2] <- "meanOfSteps"

ggplot(avgSteps, aes(interval, meanOfSteps)) + geom_line(color = "blue", size = 0.7) + labs(title = "Time Series Plot of the 5-minute Interval", x = "5-minute intervals", y = "Average Number of Steps Taken")
```
  Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
  
```{r, echo=TRUE}
avgSteps[avgSteps$meanOfSteps == max(avgSteps$meanOfSteps), ]
```


## Imputing missing values

  Calculate and report the total number of missing values in the dataset.
```{r, echo=TRUE}
sum(is.na(data))
```


  Devise a strategy for filling in all of the missing values in the dataset.
```{r, echo=TRUE}
newData <- data 
for (i in 1:nrow(newData)) {
    if (is.na(newData$steps[i])) {
        newData$steps[i] <- avgSteps[which(newData$interval[i] == avgSteps$interval), ]$meanOfSteps
    }
}

head(newData)
sum(is.na(newData))
```  

  Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r, echo=TRUE}
ggplot(newData, aes(date, steps)) + geom_bar(stat = "identity",
                                             colour = "blue", fill = "blue", width = 0.7) + facet_grid(. ~ month, scales = "free") + labs(title = "Histogram of Total Number of Steps Taken Each Day", x = "Date", y = "Total number of steps")
```  

  Mean and median total number of steps taken per day
```{r, echo=TRUE}
newTotalSteps <- aggregate(newData$steps, list(Date = newData$date), sum)$x
newMean <- mean(newTotalSteps)
newMean
newMedian <- median(newTotalSteps)
newMedian
```
  Do these values differ from the estimates from the first part of the assignment?
```{r, echo=TRUE}
oldMean <- mean(totalSteps)
oldMedian <- median(totalSteps)
newMean - oldMean
newMedian - oldMedian
```


## Are there differences in activity patterns between weekdays and weekends?
  Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r, echo=TRUE}
head(newData)
newData$weekdays <- factor(format(newData$date, "%A"))
levels(newData$weekdays)
levels(newData$weekdays) <- list(weekday = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"),
                                 weekend = c("Saturday", "Sunday"))
levels(newData$weekdays)
table(newData$weekdays)
```  
  Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r, echo=TRUE}
avgSteps <- aggregate(newData$steps, list(interval = as.numeric(as.character(newData$interval)), weekdays = newData$weekdays), mean)
names(avgSteps)[3] <- "meanOfSteps"
library(lattice)
xyplot(avgSteps$meanOfSteps ~ avgSteps$interval | avgSteps$weekdays, 
       layout = c(1, 2), type = "l", 
       xlab = "Interval", ylab = "Number of steps")
```